---
title: "correlation analysis"
author: "Feiya Suo"
date: "3/13/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Read data
```{r}
data <- read.csv("/Users/feiyasuo/Documents/GitHub/STOR765_Spring2022_FeiyaSuo/CountyLevel_data.csv")
```

# Descriptive Statistics
In this section, I calculate the descriptive statistics for each variable in the county-level data. The summary of the statistics can be found in the report.

```{r}
summary(data$available_volunteer)
sd(data$available_volunteer)

summary(data$percent_available)
sd(data$percent_available)

summary(data$volunteered)
sd(data$volunteered)

summary(data$percent_volunteered)
sd(data$percent_volunteered)

summary(data$Vaccinated_count)
sd(data$Vaccinated_count)

summary(data$Vaccinated.percent)
sd(data$Vaccinated.percent)

summary(data$percent_NP)
sd(data$percent_NP)

summary(data$percent_RN)
sd(data$percent_RN)

summary(data$percent_LPN)
sd(data$percent_LPN)

summary(data$percent_Other)
sd(data$percent_Other)

summary(data$percent_retired)
sd(data$percent_retired)

summary(data$ratio_active)
sd(data$ratio_active)
```

# Scatterplots
The following figures show the scatter plots between vaccination rates and the variables relevant to volunteer nurses. Each dot represents a datapoint, and the blue line represents the estimated regression line between the two variables. The grey area shows the 95% confidence interval of the estimation. 

```{r, fig.align="center", echo = FALSE}
data$proportion_volunteered <- data$volunteered/data$available_volunteer

library(ggplot2)
ggplot(data, aes(x=available_volunteer, y=Vaccinated.percent)) +
  geom_point() + 
  geom_text(label=data$volunteer_county, size=2, vjust=2) +
  geom_smooth(method=lm)+ 
  labs(x = "# of Available Volunteers", y = "Vaccination Coverage Rate")

ggplot(data, aes(x=percent_available, y=Vaccinated.percent)) +
  geom_point() + 
  geom_text(label=data$volunteer_county, size=2, vjust=2) +
  geom_smooth(method=lm)+ 
  labs(x = "% of Available Volunteers", y = "Vaccination Coverage Rate")

ggplot(data, aes(x=volunteered, y=Vaccinated.percent)) +
  geom_point() + 
  geom_text(label=data$volunteer_county, size=2, vjust=2) +
  geom_smooth(method=lm)+ 
  labs(x = "# of Nurses who Actually Volunteered", y = "Vaccination Coverage Rate")

ggplot(data, aes(x=percent_volunteered, y=Vaccinated.percent)) +
  geom_point() + 
  geom_text(label=data$volunteer_county, size=2, vjust=2) +
  geom_smooth(method=lm)+ 
  labs(x = "% of Nurses who Actually Volunteered", y = "Vaccination Coverage Rate")

ggplot(data, aes(x=proportion_volunteered, y=Vaccinated.percent)) +
  geom_point() + 
  geom_text(label=data$volunteer_county, size=2, vjust=2) +
  geom_smooth(method=lm)+ 
  labs(x = "Proportion of Nurses who Actually Volunteered", y = "Vaccination Coverage Rate")
```

# Hypothesis testing
The analysis here show the correlation between vaccination rate and volunteer nurses. If the p-value is smaller than 0.05, it shows the correlation is significant at 5% level. It indicates that we have strong evidence to reject the null hypothesis that the two variables are independent.

```{r}
# number of available volunteers
cor(data$Vaccinated.percent, data$available_volunteer)
m1 <- lm(data$Vaccinated.percent ~ data$percent_available)
summary(m1)

# percentage of available volunteers over county population
m1_1 <- lm(data$Vaccinated.percent ~ data$available_volunteer)
summary(m1_1)

# number of nurses who volunteered
m2 <- lm(data$Vaccinated.percent ~ data$volunteered)
summary(m2)

# percentage of nurses who volunteered over county population
m2_1 <- lm(data$Vaccinated.percent ~ data$percent_volunteered)
summary(m2_1)

# proportion of nurses who volunteered over available volunteers
m3 <- lm(data$Vaccinated.percent ~ data$proportion_volunteered)
summary(m3)

```

# Region level analysis
## aggregate data
Since there are many zeros in the county-level data, I aggregate the data to region level.

```{r}
# merge region info
region <- read.csv("/Users/feiyasuo/Documents/GitHub/STOR765_Spring2022_FeiyaSuo/region.csv")
raw_county <- read.csv("/Users/feiyasuo/Documents/GitHub/STOR765_Spring2022_FeiyaSuo/raw_data.csv")
raw_region <- merge(region, raw_county[,-1], all = TRUE, 
               by.x="County", by.y = "volunteer_county")
raw_region$Population <- as.numeric(gsub(",","",as.character(raw_region$Population)))
raw_region$Vaccinated_count <- as.numeric(gsub(",","",as.character(raw_region$Vaccinated_count)))

# aggregate data
raw_region <- raw_region[,c(2:17,19,20)]
col.nam <- colnames(raw_region)[-1]
raw_region[col.nam] <- as.data.frame(sapply(raw_region[col.nam],as.numeric))
region_data <- aggregate(. ~ AHEC.region, raw_region, sum)

# calculate metrics
final <- region_data[,c(1,10)]
final$percent_available <- region_data$available_volunteer/region_data$Population
final$volunteered <- region_data$volunteered
final$percent_volunteered <- region_data$volunteered/region_data$available_volunteer
final$Vaccinated_count <- region_data$Vaccinated_count
final$Vaccinated.percent <- region_data$Vaccinated_count/region_data$Population
final$percent_NP <- region_data$NP/region_data$available_volunteer
final$percent_RN <- region_data$RN/region_data$available_volunteer
final$percent_LPN<- region_data$LPN/region_data$available_volunteer
final$percent_NAII<- region_data$NAII/region_data$available_volunteer
final$percent_Other <- region_data$Other/region_data$available_volunteer
final$percent_retired <- region_data$Retired/region_data$available_volunteer
final$ratio_active <- region_data$Active/(region_data$Active + region_data$Inactive)

write.csv(final, "/Users/feiyasuo/Documents/GitHub/STOR765_Spring2022_FeiyaSuo/RegionLevel_data.csv")
```

## Scatterplots
The following figures show the scatter plots between vaccination rates and the variables relevant to volunteer nurses. Each dot represents a datapoint, and the blue line represents the estimated regression line between the two variables. The grey area shows the 95% confidence interval of the estimation. 

```{r}
ggplot(final, aes(x=available_volunteer, y=Vaccinated.percent)) +
  geom_point() + 
  geom_text(label=final$AHEC.region, size=2, vjust=2) +
  geom_smooth(method=lm)+ 
  labs(x = "# of Available Volunteers", y = "Vaccination Coverage Rate")

ggplot(final, aes(x=percent_available, y=Vaccinated.percent)) +
  geom_point() + 
  geom_text(label=final$AHEC.region, size=2, vjust=2) +
  geom_smooth(method=lm)+ 
  labs(x = "% of Available Volunteers", y = "Vaccination Coverage Rate")

ggplot(final, aes(x=volunteered, y=Vaccinated.percent)) +
  geom_point() + 
  geom_text(label=final$AHEC.region, size=2, vjust=2) +
  geom_smooth(method=lm)+ 
  labs(x = "# of Nurses who Actually Volunteered", y = "Vaccination Coverage Rate")

ggplot(final, aes(x=percent_volunteered, y=Vaccinated.percent)) +
  geom_point() + 
  geom_text(label=final$AHEC.region, size=2, vjust=2) +
  geom_smooth(method=lm)+ 
  labs(x = "% of Nurses who Actually Volunteered", y = "Vaccination Coverage Rate")

ggplot(final, aes(x=percent_volunteered, y=Vaccinated.percent)) +
  geom_point() + 
  geom_text(label=final$AHEC.region, size=2, vjust=2) +
  geom_smooth(method=lm)+ 
  labs(x = "Proportion of Nurses who Actually Volunteered", y = "Vaccination Coverage Rate")
```

## Hypothesis testing
The analysis here show the correlation between vaccination rate and volunteer nurses. If the p-value is smaller than 0.05, it shows the correlation is significant at 5% level. It indicates that we have strong evidence to reject the null hypothesis that the two variables are independent.

```{r}
# number of available volunteers
m1 <- lm(final$Vaccinated.percent ~ final$available_volunteer)
summary(m1)

# percentage of available volunteers over county population
m1_1 <- lm(final$Vaccinated.percent ~ final$percent_available)
summary(m1_1)

# number of nurses who volunteered
m2 <- lm(final$Vaccinated.percent ~ final$volunteered)
summary(m2)

# percentage of nurses who volunteered over county population
m2_1 <- lm(final$Vaccinated.percent ~ final$percent_volunteered)
summary(m2_1)

# proportion of nurses who volunteered over available volunteers
m3 <- lm(final$Vaccinated.percent ~ final$percent_volunteered)
summary(m3)
```


